---
title: "Case Study: Rotterdam survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: Rotterdam survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`survival::rotterdam` is a richer oncology survival dataset than `veteran`,
with tumor biology and treatment variables that allow a more structured
heterogeneity analysis.

## Objective

The goal is to study whether chemotherapy benefit varies with baseline disease
burden and receptor-related biomarkers.

At horizon \(h = 8\) years, the forest targets subgroup-specific survival
effects summarized as RMST differences or survival-probability differences,
depending on the chosen target.

## Analysis setup

```{r}
dat <- prepare_case_rotterdam()

fit <- fit_survival_forest(
  data = dat,
  time = "time",
  event = "event",
  treatment = "treatment",
  covariates = setdiff(names(dat), c("sample_id", "time", "event", "treatment")),
  sample_id = "sample_id",
  horizon = 8,
  seed = 123,
  num_trees = 400,
  tree_minbucket = 120
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

The interpretation is observational: baseline clinical and tumor features are
used to adjust treatment comparisons and to discover effect modifiers.

## Treatment and observed time

```{r}
plot_treatment_outcome(fit)
```

The treatment-outcome panel gives a first sense of follow-up scale and censoring.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

Compared with the simpler survival example, the Rotterdam case produces a more
layered subgroup structure. The estimated effects vary across receptor-related
and prognostic strata.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The explanation tree is useful here because it turns a high-dimensional survival
forest into a compact rule system involving `pgr`, `er`, `age`, and `size`.

## Variable importance

```{r}
plot_variable_importance(fit)
```

The importance ranking confirms that biologically meaningful covariates drive
most of the heterogeneity in this example.

## Interpretation

This case study is a stronger showcase for the survival workflow because:

- the covariate space is richer,
- the resulting subgroups are more structured,
- the variable-importance and tree outputs align with recognizable clinical
  axes.

## Limitations

The diagnostics for this dataset can still raise overlap or censoring warnings.
Those warnings should be interpreted seriously. They do not invalidate the
workflow, but they do limit how strongly the subgroup effects should be treated
as decision-ready estimates.
