---
title: "Case Study: card.data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: card.data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`ivmodel::card.data` is one of the standard teaching datasets for instrumental
variables. It studies returns to schooling using college proximity as an
instrument for education.

## Objective

The goal is to move from a single IV estimate to a subgroup-specific local IV
effect. The question becomes:

\[
\tau(x) = \frac{\mathrm{Cov}(Y, Z \mid X = x)}
{\mathrm{Cov}(W, Z \mid X = x)},
\]

where \(Y\) is log wage, \(W\) is years of education, \(Z\) is college
proximity, and \(X\) contains observed background variables.

## Analysis setup

```{r}
dat <- prepare_case_card_data()

fit <- fit_instrumental_forest(
  data = dat,
  outcome = "outcome",
  treatment = "treatment",
  instrument = "instrument",
  covariates = setdiff(names(dat), c("sample_id", "outcome", "treatment", "instrument")),
  sample_id = "sample_id",
  seed = 123,
  num_trees = 400,
  tree_minbucket = 160
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_instrumental_dag()
```

The DAG emphasizes why IV is needed: unmeasured determinants of schooling and
wages can bias a naive observational effect of education.

## First stage

```{r}
plot_first_stage(fit)
```

The first-stage figure shows whether the instrument visibly shifts education.
This is not a full relevance analysis, but it is a useful visual complement to
the first-stage F-statistic in `check_table`.

## Reduced form

```{r}
plot_reduced_form(fit)
```

The reduced-form plot shows whether the instrument is also associated with the
outcome. Together, the first-stage and reduced-form views help interpret the
ratio estimand used by the instrumental forest.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

The subgroup plot shows that the local IV effect is not estimated as constant.
In this example, parental education and labor-market experience help organize
the heterogeneity.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The explanation tree provides a readable summary of where the forest places its
largest local-effect differences.

## Variable importance

```{r}
plot_variable_importance(fit)
```

The top variables line up with an interpretable story: background advantage and
experience shape where education returns appear most differentiated.

## Interpretation

This case study is useful because it turns a standard IV tutorial into a
heterogeneity tutorial:

- the first stage is visible,
- the reduced form is visible,
- subgroup differences can be summarized rather than left as one global IV
  coefficient.

## Limitations

The package still reports weak-IV warnings for some leaves in this dataset.
That means subgroup interpretation should be cautious. The local IV forest is
most informative as a structured exploration of heterogeneous compliance-driven
effects, not as automatic evidence that every leaf is equally well identified.
