---
title: "Case Study: NSW job training"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: NSW job training}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`causaldata::nsw_mixtape` is a standard program-evaluation dataset built around
the National Supported Work training program. It is widely used in causal
inference tutorials because treatment is binary, outcomes are economically
meaningful, and baseline heterogeneity is plausible.

## Objective

The question here is whether the earnings benefit of job training differs by
pre-treatment labor-market history and demographic background.

The estimand is again:

\[
\tau(x) = E[Y(1) - Y(0) \mid X = x],
\]

where \(Y\) is post-program earnings and \(W\) is program participation.

## Analysis setup

```{r}
dat <- prepare_case_nsw()

fit <- fit_observational_forest(
  data = dat,
  outcome = "outcome",
  treatment = "treatment",
  covariates = setdiff(names(dat), c("sample_id", "outcome", "treatment")),
  sample_id = "sample_id",
  seed = 123,
  num_trees = 400,
  tree_minbucket = 50
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

In this tutorial, the DAG emphasizes baseline confounding by age, education,
marital status, and lagged earnings history.

## Treatment and outcome pattern

```{r}
plot_treatment_outcome(fit)
```

The raw treatment-outcome view gives the scale of the earnings outcome and
helps contextualize the subgroup-specific effect estimates.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

The subgroup summary shows a clear gradient: older and more educated strata in
this analysis tend to have larger predicted earnings gains than the youngest,
least educated stratum.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The explanation tree provides a readable program-evaluation story. Age and
education dominate the first splits, which aligns with an economic hypothesis
that labor-market readiness modifies the returns to training.

## Variable importance

```{r}
plot_variable_importance(fit)
```

Lagged earnings also appear among the leading variables, which is consistent
with the idea that pre-program attachment to the labor market helps organize
treatment heterogeneity.

## Interpretation

This case study is useful because the learned heterogeneity is substantively
readable:

- age and education anchor the subgroup tree,
- prior earnings remain important,
- estimated gains differ enough across leaves to make subgroup summaries worth
  reporting.

## Limitations

Even in a familiar teaching dataset, subgroup patterns should not be overread.
The explanation tree is still a second-stage summary of forest predictions, and
the subgrouping pattern can move under alternative tuning or alternative
covariate definitions.
