---
title: "Case Study: NHEFS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: NHEFS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`causaldata::nhefs_complete` is one of the standard epidemiologic datasets used
to study the consequences of smoking cessation. In this tutorial the outcome is
weight change, the treatment is quitting smoking, and the covariates summarize
baseline smoking behavior and demographic background.

## Objective

The goal is not only to estimate an average effect of quitting smoking on later
weight change, but to identify whether the effect is systematically larger in
some baseline strata than in others.

Formally, the target is:

\[
\tau(x) = E[Y(1) - Y(0) \mid X = x],
\]

where \(Y\) is weight change, \(W\) is smoking cessation, and \(X\) contains
baseline characteristics.

## Analysis setup

```{r}
dat <- prepare_case_nhefs()

fit <- fit_observational_forest(
  data = dat,
  outcome = "outcome",
  treatment = "treatment",
  covariates = setdiff(names(dat), c("sample_id", "outcome", "treatment")),
  sample_id = "sample_id",
  seed = 123,
  num_trees = 400,
  tree_minbucket = 100
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

The causal story is observational: baseline smoking intensity, prior weight,
and related health behaviors may influence both the decision to quit and later
weight change. The analysis therefore relies on adjustment through the baseline
covariates supplied to the forest.

## Treatment and outcome pattern

```{r}
plot_treatment_outcome(fit)
```

This figure is descriptive rather than causal. It shows how outcome values are
distributed across treatment groups before the forest has adjusted for the full
covariate set.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

The subgroup summary indicates that estimated effects are not constant. In this
analysis, the dominant split is baseline weight, with heavier participants
showing larger predicted weight gain after smoking cessation.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The explanation tree is a compact approximation of the forest’s sample-level
effect predictions. It does not replace the forest; instead, it provides a
readable subgrouping rule for reporting.

## Variable importance

```{r}
plot_variable_importance(fit)
```

The importance profile helps explain why the subgroup tree is organized around
baseline weight and smoking history variables.

## Interpretation

This case study produces a coherent epidemiologic story:

- baseline weight is the strongest modifier,
- smoking history features also contribute,
- quitting smoking is associated with larger predicted weight gain in heavier
  baseline subgroups.

The package output is therefore useful not only for estimation, but also for
describing where the treatment effect appears concentrated.

## Limitations

This remains an observational design. The results depend on:

- the adequacy of baseline adjustment,
- the absence of important unmeasured confounding,
- the stability of the subgroup rules across resampling and alternative
  covariate specifications.

The subgroup rules should therefore be treated as structured heterogeneity
summaries, not as automatically transportable decision rules.
