---
title: "Case Study: Veteran survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: Veteran survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`survival::veteran` is a compact clinical survival dataset with treatment,
follow-up time, event status, and baseline prognostic variables such as
performance status and diagnosis time.

## Objective

The main question is whether the treatment effect on survival differs across
baseline prognosis strata.

At horizon \(h = 6\) months, the analysis targets a subgroup-specific survival
effect:

\[
\tau_h(x) = E[\min(T(1), h) - \min(T(0), h) \mid X = x]
\]

under the RMST target used by the package default.

## Analysis setup

```{r}
dat <- prepare_case_veteran()

fit <- fit_survival_forest(
  data = dat,
  time = "time",
  event = "event",
  treatment = "treatment",
  covariates = setdiff(names(dat), c("sample_id", "time", "event", "treatment")),
  sample_id = "sample_id",
  horizon = 6,
  seed = 123,
  num_trees = 400,
  tree_minbucket = 25
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

For survival settings the same baseline-adjustment idea applies, but the
outcome is now right-censored time-to-event data rather than a scalar endpoint.

## Treatment and observed time

```{r}
plot_treatment_outcome(fit)
```

This figure shows the observed-time distribution by treatment, including the
event/censoring split.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

The subgroup summary suggests that treatment benefit is not uniform. Baseline
diagnosis time and Karnofsky performance status drive the main separation.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The tree shows a clinically interpretable pattern: prognosis variables organize
the treatment-effect surface, and the lowest-performance groups are separated
from better-functioning groups.

## Variable importance

```{r}
plot_variable_importance(fit)
```

The importance chart reinforces that age, diagnosis time, and Karnofsky score
carry most of the heterogeneity signal.

## Interpretation

This is a good survival tutorial because the results are interpretable:

- clinically recognizable predictors dominate,
- subgroup effects differ enough to justify forest-based heterogeneity analysis,
- the explanation tree remains readable.

## Limitations

This is still a small teaching dataset. Survival forests are sensitive to
censoring patterns, horizon choice, and treatment overlap. Subgroup summaries
should therefore be viewed as structured exploratory evidence rather than as a
finished treatment rule.
