---
title: "Case Study: SchoolingReturns"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study: SchoolingReturns}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

## Background

`ivreg::SchoolingReturns` is another returns-to-schooling dataset with college
proximity instruments and a richer collection of baseline socioeconomic
covariates than the simpler Card example.

## Objective

The purpose of this tutorial is to show the same IV workflow on a higher
dimensional baseline space. The estimand is the subgroup-specific local IV
effect of education on log wages.

## Analysis setup

```{r}
dat <- prepare_case_schooling_returns()

fit <- fit_instrumental_forest(
  data = dat,
  outcome = "outcome",
  treatment = "treatment",
  instrument = "instrument",
  covariates = setdiff(names(dat), c("sample_id", "outcome", "treatment", "instrument")),
  sample_id = "sample_id",
  seed = 123,
  num_trees = 400,
  tree_minbucket = 80
)

fit$check_table
fit$subgroup_table
```

## Design view

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_instrumental_dag()
```

The DAG is the same IV logic as in the Card example, but the baseline
covariates now allow a more detailed heterogeneity decomposition.

## First stage

```{r}
plot_first_stage(fit)
```

The first-stage panel shows the instrument-exposure link at the analysis-table
level.

## Reduced form

```{r}
plot_reduced_form(fit)
```

The reduced-form panel complements the first stage by showing the
instrument-outcome relationship.

## Heterogeneous effect summary

```{r}
plot_subgroup_effects(fit)
```

The subgroup plot suggests that local IV effects differ across baseline family
education and cognitive-score profiles.

## Explanation tree

```{r}
plot_effect_tree(fit)
```

The explanation tree turns the forest predictions into readable rules. In this
example, parental education, baseline schooling, and measured skill variables
all contribute.

## Variable importance

```{r}
plot_variable_importance(fit)
```

The importance ranking helps connect the tree back to the full forest fit.

## Interpretation

This tutorial is valuable because it shows that the instrumental workflow is
not limited to one famous teaching dataset. The same logic transfers to a
larger baseline space, while keeping a consistent set of outputs.

## Limitations

As with any local IV analysis, interpretation depends on instrument validity
and local compliance. The subgroup rules should therefore be read as summaries
of where the local IV effect appears to vary, not as universal returns to
schooling estimates for all populations.
