---
title: "Getting Started with heteff"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with heteff}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5.5)
```

```{r}
library(heteff)
```

`heteff` standardizes three generalized random forest workflows:

- observational causal forests
- causal survival forests
- instrumental forests

The package keeps one interface and one output shape across these workflows.

## Core estimands

### Observational effect

\[
\tau(x) = E[Y(1) - Y(0) \mid X = x]
\]

### Survival effect

At a fixed horizon \(h\), `heteff` estimates subgroup-specific RMST differences
or survival-probability differences.

### Instrumental effect

\[
\tau(x) = \frac{\mathrm{Cov}(Y, Z \mid X = x)}{\mathrm{Cov}(W, Z \mid X = x)}
\]

## 1. Observational workflow

```{r}
obs_data <- simulate_observational_data(n = 1600, seed = 123)

fit_obs <- fit_observational_forest(
  data = obs_data,
  outcome = "outcome",
  treatment = "treatment",
  covariates = c(
    "age", "sex", "ecog", "line", "liver_met", "tmb", "msi",
    "signature_ifn", "baseline_ctdna", "albumin", "crp", "site_volume"
  ),
  sample_id = "sample_id",
  seed = 123,
  num_trees = 400,
  tree_minbucket = 60
)

fit_obs$subgroup_table
```

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

```{r}
plot_treatment_outcome(fit_obs)
```

```{r}
plot_subgroup_effects(fit_obs)
```

```{r}
plot_effect_tree(fit_obs)
```

## 2. Survival workflow

```{r}
surv_data <- simulate_survival_data(n = 1600, seed = 321, horizon = 12)

fit_surv <- fit_survival_forest(
  data = surv_data,
  time = "time",
  event = "event",
  treatment = "treatment",
  covariates = c(
    "age", "sex", "ecog", "line", "liver_met", "inflammation",
    "biomarker", "pathway_score", "steroid"
  ),
  horizon = 12,
  sample_id = "sample_id",
  seed = 321,
  num_trees = 400,
  tree_minbucket = 60
)

fit_surv$subgroup_table
```

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_observational_dag()
```

```{r}
plot_treatment_outcome(fit_surv)
```

```{r}
plot_subgroup_effects(fit_surv)
```

```{r}
plot_effect_tree(fit_surv)
```

## 3. Instrumental workflow

```{r}
iv_data <- simulate_instrumental_data(n = 1600, seed = 456)

fit_iv <- fit_instrumental_forest(
  data = iv_data,
  outcome = "outcome",
  treatment = "treatment",
  instrument = "instrument",
  covariates = c(
    "age", "sex", "bmi", "smoking", "pc1", "pc2",
    "center", "batch", "baseline_risk"
  ),
  sample_id = "sample_id",
  seed = 456,
  num_trees = 400,
  tree_minbucket = 60
)

fit_iv$subgroup_table
```

```{r eval = requireNamespace("ggdag", quietly = TRUE) && requireNamespace("dagitty", quietly = TRUE)}
plot_instrumental_dag()
```

```{r}
plot_first_stage(fit_iv)
```

```{r}
plot_reduced_form(fit_iv)
```

```{r}
plot_subgroup_effects(fit_iv)
```

```{r}
plot_effect_tree(fit_iv)
```

## Common output tables

Every `heteff_fit` object returns the same top-level outputs:

- `effect_table`
- `subgroup_table`
- `tree_table`
- `ranking_table`
- `check_table`
- `estimand_table`
- `variable_importance`

```{r}
names(fit_iv)
```

## Export tables and plots

```{r eval = FALSE}
export_tables(fit_iv, "output")
export_plots(fit_iv, "output")
```
