---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`heteff` is a compact wrapper for three `grf` workflows:

1. observational heterogeneous effects
2. survival heterogeneous effects
3. IV-based heterogeneous local effects

## Minimal Data Shape

Use one row per sample and pass column names to each fitting function.

```{r eval = FALSE}
library(heteff)

set.seed(42)
n <- 500
df <- data.frame(
  outcome = rnorm(n),
  treatment = rbinom(n, 1, 0.5),
  instrument = rbinom(n, 1, 0.5),
  time = rexp(n, 0.25),
  status = rbinom(n, 1, 0.8),
  x1 = rnorm(n),
  x2 = rnorm(n),
  x3 = rnorm(n)
)
```

## First Fit

```{r eval = FALSE}
fit_obs <- fit_observational_forest(
  data = df,
  outcome = "outcome",
  treatment = "treatment",
  covariates = c("x1", "x2", "x3")
)

print(fit_obs)
head(as.data.frame(fit_obs))
rank_effects(fit_obs, n = 10)
plot(fit_obs)
```

## Next Steps

- read **Guides** for the estimands and assumptions
- read **Tutorials** for workflow templates
- use **Reference** for API-level details
