---
title: "Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reference}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## API Summary

| Function | Purpose |
|:--|:--|
| `fit_observational_forest()` | Heterogeneous treatment effect from observational contrast |
| `fit_survival_forest()` | Heterogeneous effect for right-censored outcomes |
| `fit_instrumental_forest()` | Heterogeneous local effect surrogate in IV setting |
| `rank_effects()` | Return top or bottom estimated effects |
| `as.data.frame()` | Extract effect table |
| `plot()` | Quick histogram of estimated effects |

## Returned Object

`heteff_fit` contains:

- `type`
- `fit` (raw `grf` forest object)
- `effect_table` (`sample`, `estimate`, `std.error`)
- `variable_importance`
- `columns` (input mapping)

## Internal Process Names

The package keeps an intentionally short internal process:

1. validate input columns
2. cast covariates to numeric matrix
3. call the selected `grf` forest
4. build a common effect table
5. return unified `heteff_fit` object

## Typical Integration Pattern

```{r eval = FALSE}
fit <- fit_observational_forest(df, "outcome", "treatment", c("x1", "x2"))
tab <- as.data.frame(fit)
top <- rank_effects(fit, n = 25)
```
